A simulation-based assessment of the relation between Stone Age sites and relative sea-level change along the Norwegian Skagerrak coast

### 2022-03-11 analysis/paper/paper.Rmd
Trying to deal with multiple phases at each site

```{r shoredate, echo = FALSE, fig.cap = "The result of backwards shoreline dating the sites with radiocarbon dates corresponding to the artefact inventory using the method proposed here. The shoreline dates are plotted in grey and underlined with the 95\\% HDR in black. These are plotted against the modelled radiocarbon dates for each site, which are given colour from oldest to youngest phase.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/shoredate.png"))
```

```{r shoredatediff, echo = FALSE, fig.cap = "The agreement between the shoreline dates and radiocarbon dates given in Figure ref(fig:shoredate) are evaluated using the method described by Parnell et al. (2008). When the range of the 95\\% HDRs for age difference crosses zero the dates are considered synchronous. The colour coding follows the divison of phases given in Figure ref(fig:shoredate).", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/shoredate2.png"))
```

@@

```{r redate, echo = FALSE, fig.cap = "Re-dating previously shoreline dated sites in the study area without radiocarbon dates or with radiocarbon dates that do not correspond to the artefact inventories. The range of the 95\\% HDR in grey are compared to the dates originally proposed by the excavation reports in red.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/redate.png"))
```

@@

```{r brunlanes, echo = FALSE, fig.cap = "Shoreline dating of the Brunlanes sites using site altitudes provided by Jaksland (2014:tab.4). A) The result of applying the approach to shoreline dating outlined above. Dates provided by Jaksland (2014) are plotted in red. The box indicates a 50 year uncertainty range which in combination with the red line extends 200 years. Follwing Jaksland's approach these start from the earliest possible date. B) Map showing the centroids of the Pauler sites and Sky 1. The sea-level has been simulated using the probability density associated with the shoreline date for Pauler 1 (see also map in Jaksland 2014:fig.12a). Pauler 1 is the red point.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/brunlanes.png"))
```

@@

It is worth highlighting some sources of likely variation and uncertainty that have not been considered here. First of all the DTM has only been corrected for major modern disturbances. This means that erosion, although likely not that prevalent, has not been taken into account. Secondly, the DTM has a vertical error [cf. @lewis2021]. Thirdly, the displacement curves were here interpolated to all site locations without accounting for increased uncertainty as one moves futher away from the isobases of the displacement curves. Finally, the question of how site limits are defined nor the elevation range over which these extend was given much consideration. While each of these factors will have variable but likely not too large of an impact on the results, they clearly represent dimensions which would all benefit from further exploration.

@@

Finally, this analysis was based on a simulation approach for integrating multiple sources of spatio-temporal uncertainty. Here this was simply used to inform the question of the distance between sites and the shore-line. However, this method and general framework can be extended to a wide range of use-cases where one needs to visualise and quantitatively or qualitatively evaluate the relationship between archaeological phenomena, the prehistoric shore-line, and the uncertainty inherent in this reconstruction.

>>>>

```{r shoredate, echo = FALSE, fig.cap = "The result of backwards shoreline dating the sites with radiocarbon dates corresponding to the artefact inventory using the method proposed here. The shoreline dates are plotted in grey and underlined with the 95\\% HDR in black. These are plotted against the modelled radiocarbon dates which are given colour from oldest to youngest phase for each site.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/shoredate.png"))
```

```{r shoredatediff, echo = FALSE, fig.cap = "Evaluation of the agreement between the shoreline dates and radiocarbon dates given in Figure ref(fig:shoredate), based on the method described by Parnell et al. (2008). When the range of the 95\\% HDRs for age difference crosses zero and are coloured black, the shoreline- and radiocarbon dates are considered synchronous. The colour coding follows the divison of phases given in Figure ref(fig:shoredate).", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/shoredate2.png"))
```

@@

```{r redate, echo = FALSE, fig.cap = "Re-dating previously shoreline dated sites in the study area without radiocarbon dates or with radiocarbon dates that do not correspond to the artefact inventories. The range of the 95\\% HDRs in grey are compared to the dates originally proposed by the excavation reports in red.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/redate.png"))
```

@@

```{r brunlanes, echo = FALSE, fig.cap = "Shoreline dating of the Brunlanes sites using site altitudes provided by Jaksland (2014:tab.4). A) The result of applying the approach to shoreline dating outlined above. The shoreline date in grey is underlined with the 95\\% HDR in black. Dates provided by Jaksland (2014) are plotted in red. Follwing his approach, the box indicates a 50 year uncertainty range which in combination with the red line extends 200 years. B) Map showing the centroids of the Pauler sites and Sky 1. The sea-level has been simulated using the probability density associated with the shoreline date for Pauler 1 (see also map in Jaksland 2014:fig.12a). Pauler 1 is the red point.", out.width = "525px", fig.align="center", regfloat=TRUE}
knitr::include_graphics(here("analysis/figures/brunlanes.png"))
```

@@

It is worth highlighting some limitations and sources of likely variation and uncertainty that have not been considered here. First of all the sample size is quite strained, and the future addition of only a few sites might skew the picture significantly. Secondly, the DTM has only been corrected for major modern disturbances. This means that erosion, although likely not that prevalent, has not been taken into account. Thirdly, the DTM has a vertical error of, which could also benefit from being integrated in the analysis [cf. @lewis2021]. Fourthly, the displacement curves were here interpolated to all site locations without accounting for increased uncertainty as one moves further away from the isobases of the displacement curves. Fifthly, neither the question of how site limits are defined nor the elevation range over which these extend was given much consideration. Finally, the division of phases at each sites was here simply done by treating  While each of these factors will have variable impact on the final results, they clearly represent dimensions which would all benefit from further consideration.

@@

Finally, this analysis employed a simulation approach to integrate multiple sources of spatio-temporal uncertainty. Here this was simply used to inform the question of the distance between sites and the shore-line. However, this method and general framework can be extended to a wide range of use-cases where one needs to visualise, and quantitatively or qualitatively evaluate the relationship between archaeological phenomena, the prehistoric shore-line, and the uncertainty inherent in this reconstruction.
